The function--sheaf dictionary for smooth commutative group schemes over finite fields

### 2014-08-24 crelle-dictionary.tex
Fixed most of the spacing issues.

Now suppose $\qcs{L} = (\gqcs{L},\mu,\phi)$ is a quasicharacter sheaf on $G$.
The isomorphism $\mu : \bm^* \gqcs{L} \to \gqcs{L} \boxtimes\gqcs{L} $ together with
\ref{QC.3} guarantee
that the function $\trFrob{(\gqcs{L},\phi)} : G(\Fq)\to \EEx$ is a group homomorphism, which we will also denote by $\trFrob{\qcs{L}}$.  Moreover,
this homomorphism depends only on the isomorphism class of $\qcs{L}$, so we obtain a map
\begin{align*}
\TrFrob{G} : \QCiso{G} &\to G(\Fq)^* \\
\qcs{L} &\mapsto \trFrob{\qcs{L}}.
\end{align*}
Since tensor products on the stalks of $\qcs{L}$ induce pointwise multiplication on the trace of Frobenius, $\TrFrob{G}$ is a group homomorphism.  We devote the remainder of this section to
studying the behavior of the trace of Frobenius under morphisms of group schemes.  These simple but important
results on pullbacks and external products will be used in the proofs of many results in this paper.

@@

Now, let $\qcs{L} \ceq (\gqcs{L},\mu,\phi)$ be a quasicharacter sheaf on $G$.
Set $\qcs{L}_1 \ceq q_1^* \qcs{L}$ and $\qcs{L}_2 \ceq q_2^* \qcs{L}$;
by Lemma~\ref{lem:pullback}, $\qcs{L}_1$ a quasicharacter sheaf on $G_1$
and $\qcs{L}_2$ is a quasicharacter sheaf on $G_2$.
We will obtain an isomorphism $\qcs{L} \iso  \qcs{L}_1\boxtimes \qcs{L}_2$.
Applying the functor $f^*$ to the isomorphism $\mu$ yields
\begin{equation}\label{eq:fm}
f^*\mu : f^* m^* \gqcs{L} \to f^*(\gqcs{L}\boxtimes \gqcs{L}) .
\end{equation}
We have already observed that $m\circ f = \id_G$, so $f^* m^* \gqcs{L} = \gqcs{L}$.
On the other hand, $f^*(\gqcs{L}\boxtimes \gqcs{L}) = f^*p_1^*\gqcs{L}\otimes f^* p_2^*\gqcs{L}$
by definition.  Since $f^*p_1^*\gqcs{L} = r_1^* q_1^* \gqcs{L}$ and $f^*p_2^*\gqcs{L} = r_2^* q_2^* \gqcs{L}$
we get that $f^*(\gqcs{L}\boxtimes \gqcs{L}) = \gqcs{L}_1\boxtimes \gqcs{L}_2$.
It follows that \eqref{eq:fm} gives an isomorphism $\gqcs{L} \to  \gqcs{L}_1\boxtimes \gqcs{L}_2$.
It is a matter of routine to show that this morphism satisfies
\ref{QC.3} as it applies here,
from which it follows that we have exhibited an isomorphism
$\qcs{L} \to \qcs{L}_1\boxtimes \qcs{L}_2$ of quasicharacters sheaves on $G\times G$.
This completes the proof that $\boxtimes$ is essentially surjective.
\end{proof}

@@

\begin{proposition}
If $G$ is \'etale then $S_G$ is an isomorphism.
\end{proposition}
\begin{proof}
We first show that $S_G$ is injective.  Suppose $\qcs{L} \in \QCiso{G}$ with $S_G(\qcs{L}) = [\alpha \oplus \beta] = 0$,
so that $\alpha \oplus \beta = d\rho$ for some $\rho \in C^0(\Weil{},C^1({\bar G},\EEx) = C^1({\bar G},\EEx)$.
For each $x\in {\bar G}$, define $\rho_x : \gqcs{L}_x \to \EEx$ by $v_x \mapsto \rho(x)$.
Then the indexed set of isomorphisms $\{ \gamma_x : \gqcs{L}_x \to \EE \tq x\in {\bar G}\}$
defines an isomorphism $\qcs{L} \to (\EE)_G$.
Since $\qcs{L} = 0 \in \QCiso{G}$, $S_G$ is injective.

@@

\begin{proposition}\label{prop:BG}
Extension of scalars defines a full and faithful functor $B_G : \bQC(G) \to \QC(G)$.
\end{proposition}

@@

\begin{lemma}\label{lem:bounded_connected}
If $G$ is a connected commutative algebraic group over $\Fq$ then $B_G : \bQC(G) \to \QC(G)$ is an equivalence of categories.
\end{lemma}

@@

  Since $b_G^*$ is full and faithful
  \citelist{\cite{deligne-katz:SGA7.2}*{Expos\'e XIII} \cite{beilinson-bernstein-deligne:81a}*{Prop. 5.2.1}},
  $
  b_{G\times G}^* : \Hom(m^*\qcs{E},\qcs{E}\boxtimes\qcs{E}) \to \Hom({\bar m}^*\gqcs{L},\gqcs{L}\boxtimes\gqcs{L})
  $
  is a bijection
  (hom taken in the categories on constructible $\ell$-adic sheaves on
  $G\times G$ and ${\bar G}\times {\bar G}$ respectively,
  in which $\ell$-adic local systems sit as full subcategories).
  Let $\mu_\qcs{E} : m^*\qcs{E} \to \qcs{E}\boxtimes\qcs{E}$ be the isomorphism matching
  $\mu : {\bar m}^*\gqcs{L} \to \gqcs{L}\boxtimes\gqcs{L}$,
  the latter appearing in the definition of $\qcs{L}$.
  Then, as in Section~\ref{ssec:descentG}, $(\qcs{E},\mu_\qcs{E})$ is an object in $\bQC(G)$
  and $\qcs{L} \ceq (\gqcs{L},\mu,\phi)$ is isomorphic to $(b_G^*\qcs{E},b_{G\times G}^*\mu_\qcs{E})$ in $\QC(G)$.
  Thus, the full and faithful functor $B_G : \bQC(G) \to \QC(G)$ from Section~\ref{ssec:descentG}
  is also essentially surjective, hence an equivalence.
\end{proof}

@@

  Extensions of $G^0(\bFq)$ by $A(\bFq)$, such as $B(\bFq)$,
  correspond to classes in $\Ext^1_{\ZZ[\Weil{}]}(G^0(\bFq), A(\bFq))$.
  Similarly, extensions of $G(\bFq)$ by $A(\bFq)$ correspond to
  classes in $\Ext^1_{\ZZ[\Weil{}]}(G(\bFq), A(\bFq))$.  The map
  $G^0(\bFq) \to G(\bFq)$ induces the map
  \[
  \Ext^1_{\ZZ[\Weil{}]}(G(\bFq), A(\bFq)) \to \Ext^1_{\ZZ[\Weil{}]}(G^0(\bFq), A(\bFq))
  \]
  fitting into the long exact sequence derived from applying
  the functor $\Hom(\mbox{---}, A(\bFq))$ to $G^0(\bFq) \to G(\bFq) \to \pi_0(G)(\bFq)$:
  \[
  \Ext^1_{\ZZ[\Weil{}]}(G(\bFq), A(\bFq)) \to \Ext^1_{\ZZ[\Weil{}]}(G^0(\bFq), A(\bFq)) \to \Ext^2_{\ZZ[\Weil{}]}(\pi_0(G)(\bFq), A(\bFq)).
  \]
  Since $\Weil{} \cong \ZZ$ has cohomological dimension $1$ \cite{brown:CohomologyGrps}*{Ex. 4.3},
  $\Ext^2_{\ZZ[\Weil{}]}(\pi_0(G)(\bFq), A(\bFq))$ vanishes \cite{cartan-eilenberg:HomologicalAlgebra}*{Thm. 2.6}.

@@

\begin{proof}
  Recall the short exact sequence \eqref{eq:pi0}
  defining the component group scheme for $G$:
  \[
  \begin{tikzcd}
    1 \rar & G^0 \rar & G \rar & \pi_0(G) \rar & 0.
  \end{tikzcd}
  \]
  Let
  \[
  \begin{tikzcd}[row sep=20]
  & 0 \arrow{d} & 0 \arrow{d} & 0\arrow{d} \\
    0 \arrow[dashed]{r} & \ker \TrFrob{\pi_0(G)} \arrow[dashed]{r} \dar & \arrow{d} \ker \TrFrob{G} \arrow[dashed]{r} & \ker \TrFrob{G^0} =0 \arrow[dotted, out=-10, in=170]{dddll} \dar & \\
    0 \rar & \QCiso{\pi_0(G)} \rar \dar{\TrFrob{\pi_0(G)}}
    & \QCiso{G} \rar \dar{\TrFrob{G}} & \QCiso{G^0} \rar \dar{\TrFrob{G^0}} & 0\\
    0 \rar & \pi_0(G)(\Fq)^* \rar \dar
    & \arrow{d} G(\Fq)^* \rar & G^0(\Fq)^* \rar \dar & 0\\
    & \arrow{d} \coker \TrFrob{\pi_0(G)} =0 \arrow[dashed]{r} & \arrow{d} \coker \TrFrob{G} \arrow[dashed]{r} &  \arrow{d} \coker \TrFrob{G^0} =0 & \\
      & 0 & 0 & 0 \\
  \end{tikzcd}
  \]
  be the commutative diagram of abelian groups obtained by applying
  Lemma~\ref{lem:pullback} to \eqref{eq:pi0}.
 %
  The sequence of abelian groups
  \begin{equation}\label{eq:pi0k}
  \begin{tikzcd}
    1 \rar & G^0(\Fq) \rar & G(\Fq) \rar & \pi_0(G)(\Fq) \rar & 0,
  \end{tikzcd}
  \end{equation}
  is exact since $\Hh^1(\Fq,G^0) =0$ by Lemma~\ref{lem:G0alg-grp} and Lang's theorem on connected algebraic groups over finite fields \cite{lang:56a}.
  Since $\EEx$ is divisible, $\Hom_\text{Ab}(\ - \ ,\EEx)$ is exact and thus the dual sequence of
  character groups is also exact.
%
  The upper row is exact by Proposition~\ref{prop:middleexact}.
  Now Lemma~\ref{lem:G0alg-grp} and Proposition~\ref{prop:connected}
  imply that $\ker \TrFrob{G^0} =0$ and $\coker \TrFrob{G^0}=0$,
  while Proposition~\ref{prop:etale-iso} gives $\ker \TrFrob{\pi_0(G)} \iso \Hh^0(\Weil{},\Hh^2(\pi_0({\bar G}),\EEx))$
  and $\coker \TrFrob{\pi_0(G)}=0$.
%
It follows from the snake lemma that $\coker \TrFrob{G} =0$
and $\ker \TrFrob{\pi_0(G)} \to \ker \TrFrob{G}$ is an isomorphism.
This calculates $\ker  \TrFrob{G}$ and shows that
\[
  \begin{tikzcd}
0 \arrow{r} & \Hh^0(\Weil{},\Hh^2(\pi_0(\bG),\EEx)) \arrow{r} & \QCiso{G} \arrow{r}{\TrFrob{G}} & G(\Fq)^* \arrow{r} & 0 
  \end{tikzcd}
\]
is exact.
\end{proof}

@@

\begin{proposition} \label{prop:csbe}
Set $G_{k'} \ceq G\times_\Spec{k} \Spec{k'}$ and let
\[
\QC(\Res_{k'/k}(G_{k'})) \xrightarrow{\iota^*} \QC(G)
\]
be the functor obtained by pullback along the canonical closed immersion of $k$-schemes
$\iota : G \hookrightarrow \Res_{k'/k}(G_{k'})$
\cite{bosch-lutkebohmert-reynaud:NeronModels}*{\S 7.6}.
The following diagram commutes:
\[
\begin{tikzcd}
\QCiso{\Res_{k'/k}(G_{k'})} \arrow[two heads]{r}{\iota^*} \dar[swap]{\TrFrob{\Res_{k'/k}(G_{k'})}} & \QCiso{G} \dar{\TrFrob{G}} \\
G(k')^* \arrow[two heads]{r}{i_{k'/k}^*} & G(k)^*.
\end{tikzcd}
\]
\end{proposition}
\begin{proof}
This result follows immediately from Lemma~\ref{lem:pullback} together with the identifications
\[
\Res_{k'/k}(G_{k'})(k) \cong G_{k'}(k') \cong G(k')
\]
from the definitions of Weil restriction and base change.
\end{proof}

@@

If $\qcs{L} \ceq (\gqcs{L}, \mu, \phi)$ is a quasicharacter sheaf on $G$, we define
a quasicharacter sheaf $\qcs{L}' \ceq (\gqcs{L}, \mu, \phi_{k'})$ on the base change
$G_{k'}$ of $G$ to $k'$ by setting
\[
\phi_{k'} \ceq \phi \circ \Frob{G}^*(\phi) \circ \cdots \circ (\Frob{G}^{n-1})^*(\phi).
\]
The commutativity of the diagram (QC.3) for $\phi_{k'}$
follows from the fact that $\Frob{G_{k'}} = \Frob{G}^n$.
Note that we may also think about the construction of $\phi_{k'}$ from $\phi$
as taking the action $\varphi$ of $\Weil{k}$ on $\gqcs{L}$
defined in Section~\ref{sec:category} and restricting it to the subgroup $\Weil{k'}$.

@@

\begin{proof}
Let $\qcs{L} \ceq (\gqcs{L}, \mu, \phi) \in \QC(G)$ and write $F$ for $\Frob{G}$.  For any $x \in G(k')$,
we may compute the value of $t_{G_{k'}}(\nu_{k'/k}\qcs{L})(x)= t_{\nu_{k'/k}\qcs{L}}(x)$ as the trace of $\phi_{k'}$ on $\gqcs{L}_x$,
and the value of $\Nm^*(\TrFrob{G}(\qcs{L}))(x)$ as the trace of $\phi$ on $\gqcs{L}_{\Nm(x)}$.
Applying \ref{QC.3} to the stalk of $\gqcs{L}^{\boxtimes n}$ at the point $(x, \Frob{}(x), \ldots, \Frob{}^{n-1}(x))$ yields a diagram
\[
\begin{tikzcd}
\gqcs{L}_{\Nm(x)} \rar \dar{\phi_{\Nm(x)}} & \gqcs{L}_{F(x)} \otimes \gqcs{L}_{F^2(x)} \otimes \cdots \otimes \gqcs{L}_x
\dar{\phi_x \otimes (F^*\phi)_x \otimes \cdots \otimes ((F^{n-1})^*\phi)_x} \\
\gqcs{L}_{\Nm(x)} \rar & \gqcs{L}_x \otimes \gqcs{L}_{F(x)} \otimes \cdots \otimes \gqcs{L}_{F^{n-1}(x)}.
\end{tikzcd}
\]
Choose a basis vector $v$ for $\gqcs{L}_{\Nm(x)}$ and write the image of $v$ under the
bottom map as $v_0 \otimes v_1 \otimes \cdots \otimes v_{n-1}$
for $v_i \in \gqcs{L}_{\Frob{}^i(x)}$.  By \ref{QC.2}, $v$ maps to
$v_1 \otimes v_2 \otimes \cdots \otimes v_0$ along the top of the diagram.
Let $\alpha_i \in \EEx$ represent $((F^i)^*\phi)_x$ with respect to these bases and let $\alpha$ be
the trace of $\phi_{\Nm(x)}$.  We may now equate the trace $\alpha$ of $\phi$ on $\gqcs{L}_{\Nm(x)}$
with the product $\alpha_0 \cdots \alpha_{n-1}$, which is the trace of $\phi_{k'}$ on $\gqcs{L}_x$.
\end{proof}

@@

Let $K$, $R$ and $R_n$ be as above.
Set $S = \Spec{R}$ and $S_n = \Spec{R_n}$, for every non-negative integer $n$;
note that $S_0 = \Spec{\Fq}$ is the special fibre of $S$.

We will need the Greenberg functor, which provides
a functor from the category of schemes 
over $R_n$, for each $n\in \NN$, to the category of schemes over $\Fq$;
See \cite{bertrapelle-gonzales:Greenberg} for the definition and fundamental properties of the Greenberg functor as we use it; other important references include
\citelist{
\cite{greenberg:61}, \cite{greenberg:63a},
\cite{demazure-gabriel:GroupesAlgebriques}*{V, \S 4, no. 1},
\cite{bosch-lutkebohmert-reynaud:NeronModels}*{Ch. 9, \S 6},
\cite{nicaise-sebag:motivicSerre}*{\S 2.2}
}.
This functor respects open immersions, closed immersions, \'etale morphisms, smooth morphisms and geometric components.
%
For any non-negative integer $n$ we will write
\[
\Gr^R_n : \partop{\text{Schemes}}{\text{lft over $R$}} \to \partop{\text{Schemes}}{\text{lft over $\Fq$}}
\]
for the functor produced by composing pullback along $S_n \to S$ with the Greenberg functor. 
This functor also respects open immersions, closed immersions, \'etale morphisms, smooth morphisms and geometric components, and
\[
\Gr^R_n(X)(\Fq) \iso X(R_n),
\]
canonically, for any $X$ lft over $S$.

>>>>

Now suppose $\qcs{L} = (\gqcs{L},\mu,\phi)$ is a quasicharacter sheaf on $G$.
Then $\bm^* \gqcs{L} \iso \gqcs{L} \boxtimes\gqcs{L} $ together with
\ref{QC.3} guarantee
that the function $\trFrob{(\gqcs{L},\phi)} : G(\Fq)\to \EEx$ is a group homomorphism, which we will also denote by $\trFrob{\qcs{L}}$.  Moreover,
this homomorphism depends only on the isomorphism class of $\qcs{L}$, so we obtain a map
\begin{align*}
\TrFrob{G} : \QCiso{G} &\to G(\Fq)^* \\
\qcs{L} &\mapsto \trFrob{\qcs{L}}.
\end{align*}
Since tensor products on the stalks of $\qcs{L}$ induce pointwise multiplication on the trace of Frobenius, $\TrFrob{G}$ is a group homomorphism.  We devote the remainder of this section to
studying the behavior of the trace of Frobenius under morphisms of group schemes.  These simple but important
results on pullbacks and external products will be used in the proofs of many results in this paper.

@@

Now, let $\qcs{L} \ceq (\gqcs{L},\mu,\phi)$ be a quasicharacter sheaf on $G$.
Set $\qcs{L}_1 \ceq q_1^* \qcs{L}$ and $\qcs{L}_2 \ceq q_2^* \qcs{L}$;
by Lemma~\ref{lem:pullback}, $\qcs{L}_1$ a quasicharacter sheaf on $G_1$
and $\qcs{L}_2$ is a quasicharacter sheaf on $G_2$.
We will obtain an isomorphism $\qcs{L} \iso  \qcs{L}_1\boxtimes \qcs{L}_2$.
Applying the functor $f^*$ to the isomorphism $\mu$ yields
\begin{equation}\label{eq:fm}
f^*\mu : f^* m^* \gqcs{L} \to f^*(\gqcs{L}\boxtimes \gqcs{L}) .
\end{equation}
We have already seen that $m\circ f = \id_G$, so $f^* m^* \gqcs{L} = \gqcs{L}$.  
Since $f^*p_1^*\gqcs{L} = r_1^* q_1^* \gqcs{L} = r_1^* \gqcs{L}_1$ and $f^*p_2^*\gqcs{L} = r_2^* q_2^* \gqcs{L} = r_2^*\gqcs{L}_2$,
we have 
\[
f^*(\gqcs{L}\boxtimes \gqcs{L})  = f^*p_1^*\gqcs{L}\otimes f^* p_2^*\gqcs{L} = \gqcs{L}_1\boxtimes \gqcs{L}_2.
\]
It follows that \eqref{eq:fm} gives an isomorphism $\gqcs{L} \to  \gqcs{L}_1\boxtimes \gqcs{L}_2$.
It is a matter of routine to show that this morphism satisfies
\ref{QC.3} as it applies here,
from which it follows that we have exhibited an isomorphism
$\qcs{L} \to \qcs{L}_1\boxtimes \qcs{L}_2$ of quasicharacters sheaves on $G\times G$.
This completes the proof that $\boxtimes$ is essentially surjective.
\end{proof}

@@

\begin{proposition}
If $G$ is \'etale then $S_G$ is an isomorphism.
\end{proposition}
\begin{proof}
%We first show that $S_G$ is injective.  
Suppose $\qcs{L} \in \QCiso{G}$ with $S_G(\qcs{L}) = [\alpha \oplus \beta] = 0$,
so that $\alpha \oplus \beta = d\rho$ for some $\rho \in C^0(\Weil{},C^1({\bar G},\EEx) = C^1({\bar G},\EEx)$.
For each $x\in {\bar G}$, define $\rho_x : \gqcs{L}_x \to \EEx$ by $v_x \mapsto \rho(x)$.
Then the indexed set of isomorphisms $\{ \gamma_x : \gqcs{L}_x \to \EE \tq x\in {\bar G}\}$
defines an isomorphism $\qcs{L} \to (\EE)_G$.
Since $\qcs{L} = 0 \in \QCiso{G}$, $S_G$ is injective.

@@

\begin{proposition}\label{prop:BG}
Extension of scalars defines a full and faithful functor
\[
B_G : \bQC(G) \to \QC(G).
\]
\end{proposition}

@@

\begin{lemma}\label{lem:bounded_connected}
If $G$ is a connected commutative algebraic group over $\Fq$ then 
\[
B_G : \bQC(G) \to \QC(G)
\]
 is an equivalence of categories.
\end{lemma}

@@

  Since $b_G^*$ is full and faithful
  \citelist{\cite{deligne-katz:SGA7.2}*{Expos\'e XIII} \cite{beilinson-bernstein-deligne:81a}*{Prop. 5.2.1}},
 \[
  b_{G\times G}^* : \Hom(m^*\qcs{E},\qcs{E}\boxtimes\qcs{E}) \to \Hom({\bar m}^*\gqcs{L},\gqcs{L}\boxtimes\gqcs{L})
 \]
  is a bijection
  (hom taken in the categories on constructible $\ell$-adic sheaves on
  $G\times G$ and ${\bar G}\times {\bar G}$ respectively,
  in which $\ell$-adic local systems sit as full subcategories).
  Let $\mu_\qcs{E} : m^*\qcs{E} \to \qcs{E}\boxtimes\qcs{E}$ be the isomorphism matching
  $\mu : {\bar m}^*\gqcs{L} \to \gqcs{L}\boxtimes\gqcs{L}$,
  the latter appearing in the definition of $\qcs{L}$.
  Then, as in Section~\ref{ssec:descentG}, $(\qcs{E},\mu_\qcs{E})$ is an object in $\bQC(G)$
  and $\qcs{L} \ceq (\gqcs{L},\mu,\phi)$ is isomorphic to $(b_G^*\qcs{E},b_{G\times G}^*\mu_\qcs{E})$ in $\QC(G)$.
  Thus, the full and faithful functor $B_G : \bQC(G) \to \QC(G)$ from Section~\ref{ssec:descentG}
  is also essentially surjective, hence an equivalence.
\end{proof}

@@

  Extensions of $G^0(\bFq)$ by $A(\bFq)$, such as $B(\bFq)$,
  correspond to classes in $\Ext^1_{\ZZ[\Weil{}]}(G^0(\bFq), A(\bFq))$.
  Similarly, extensions of $G(\bFq)$ by $A(\bFq)$ correspond to
  classes in $\Ext^1_{\ZZ[\Weil{}]}(G(\bFq), A(\bFq))$.  The map
  $G^0(\bFq) \to G(\bFq)$ induces the map
  \[
  \Ext^1_{\ZZ[\Weil{}]}(G(\bFq), A(\bFq)) \to \Ext^1_{\ZZ[\Weil{}]}(G^0(\bFq), A(\bFq))
  \]
  fitting into the long exact sequence 
  \[
  \Ext^1_{\ZZ[\Weil{}]}(G(\bFq), A(\bFq)) \to \Ext^1_{\ZZ[\Weil{}]}(G^0(\bFq), A(\bFq)) \to \Ext^2_{\ZZ[\Weil{}]}(\pi_0(G)(\bFq), A(\bFq))
  \]
  derived from applying
  the functor $\Hom(\mbox{---}, A(\bFq))$ to $G^0(\bFq) \to G(\bFq) \to \pi_0(G)(\bFq)$.
  Since $\Weil{} \cong \ZZ$ has cohomological dimension $1$ \cite{brown:CohomologyGrps}*{Ex. 4.3},
  $\Ext^2_{\ZZ[\Weil{}]}(\pi_0(G)(\bFq), A(\bFq))$ vanishes \cite{cartan-eilenberg:HomologicalAlgebra}*{Thm. 2.6}.

@@

\begin{proof}
  Recall the short exact sequence \eqref{eq:pi0}
  defining the component group scheme for $G$:
  \[
  \begin{tikzcd}
    1 \rar & G^0 \rar & G \rar & \pi_0(G) \rar & 0.
  \end{tikzcd}
  \]
  Let
  \[
  \begin{tikzcd}[row sep=20]
%  & 0 \arrow{d} & 0 \arrow{d} & 0\arrow{d} \\
    0 \arrow[dashed]{r} & \ker \TrFrob{\pi_0(G)} \arrow[dashed]{r} \dar & \arrow{d} \ker \TrFrob{G} \arrow[dashed]{r} & \ker \TrFrob{G^0} =0 \arrow[dotted, out=-10, in=170]{dddll} \dar & \\
    0 \rar & \QCiso{\pi_0(G)} \rar \dar{\TrFrob{\pi_0(G)}}
    & \QCiso{G} \rar \dar{\TrFrob{G}} & \QCiso{G^0} \rar \dar{\TrFrob{G^0}} & 0\\
    0 \rar & \pi_0(G)(\Fq)^* \rar \dar
    & \arrow{d} G(\Fq)^* \rar & G^0(\Fq)^* \rar \dar & 0\\
   &  \coker \TrFrob{\pi_0(G)} =0 \arrow[dashed]{r} & \coker \TrFrob{G} \arrow[dashed]{r} & \coker \TrFrob{G^0} =0 & 
%    & \arrow{d} \coker \TrFrob{\pi_0(G)} =0 \arrow[dashed]{r} & \arrow{d} \coker \TrFrob{G} \arrow[dashed]{r} &  \arrow{d} \coker \TrFrob{G^0} =0 & \\
%      & 0 & 0 & 0 \\
  \end{tikzcd}
  \]
  be the commutative diagram of abelian groups obtained by applying
  Lemma~\ref{lem:pullback} to \eqref{eq:pi0}.
 %
  The sequence of abelian groups
  \begin{equation}\label{eq:pi0k}
  \begin{tikzcd}
    1 \rar & G^0(\Fq) \rar & G(\Fq) \rar & \pi_0(G)(\Fq) \rar & 0,
  \end{tikzcd}
  \end{equation}
  is exact since $\Hh^1(\Fq,G^0) =0$ by Lemma~\ref{lem:G0alg-grp} and Lang's theorem on connected algebraic groups over finite fields \cite{lang:56a}.
  Since $\EEx$ is divisible, $\Hom_\text{Ab}(\ - \ ,\EEx)$ is exact and thus the dual sequence of
  character groups is also exact.
%
  The upper row is exact by Proposition~\ref{prop:middleexact}.
  Now Lemma~\ref{lem:G0alg-grp} and Proposition~\ref{prop:connected}
  imply that $\ker \TrFrob{G^0} =0$ and $\coker \TrFrob{G^0}=0$,
  while Proposition~\ref{prop:etale-iso} gives $\ker \TrFrob{\pi_0(G)} \iso \Hh^0(\Weil{},\Hh^2(\pi_0({\bar G}),\EEx))$
  and $\coker \TrFrob{\pi_0(G)}=0$.
%
It follows from the snake lemma that $\coker \TrFrob{G} =0$
and $\ker \TrFrob{\pi_0(G)} \to \ker \TrFrob{G}$ is an isomorphism.
This calculates $\ker  \TrFrob{G}$ and shows that
\[
  \begin{tikzcd}
0 \arrow{r} & \Hh^0(\Weil{},\Hh^2(\pi_0(\bG),\EEx)) \arrow{r} & \QCiso{G} \arrow{r}{\TrFrob{G}} & G(\Fq)^* \arrow{r} & 0 
  \end{tikzcd}
\]
is exact.
\end{proof}

@@

\begin{proposition} \label{prop:csbe}
Set $G_{k'} \ceq G\times_\Spec{k} \Spec{k'}$ and let
\[
\QC(\Res_{k'/k}(G_{k'})) \xrightarrow{\iota^*} \QC(G)
\]
be the functor obtained by pullback along the canonical closed immersion 
$\iota : G \hookrightarrow \Res_{k'/k}(G_{k'})$ of $k$-schemes
given by \cite{bosch-lutkebohmert-reynaud:NeronModels}*{\S 7.6}.
The following diagram commutes:
\[
\begin{tikzcd}
\QCiso{\Res_{k'/k}(G_{k'})} \arrow[two heads]{r}{\iota^*} \dar[swap]{\TrFrob{\Res_{k'/k}(G_{k'})}} & \QCiso{G} \dar{\TrFrob{G}} \\
G(k')^* \arrow[two heads]{r}{i_{k'/k}^*} & G(k)^*.
\end{tikzcd}
\]
\end{proposition}
\begin{proof}
This result follows immediately from Lemma~\ref{lem:pullback} together with the identifications
\[
\Res_{k'/k}(G_{k'})(k) \cong G_{k'}(k') \cong G(k')
\]
from the definitions of Weil restriction and base change.
\end{proof}

@@

If $\qcs{L} \ceq (\gqcs{L}, \mu, \phi)$ is a quasicharacter sheaf on $G$, we define
%a quasicharacter sheaf 
$\qcs{L}' \ceq (\gqcs{L}, \mu, \phi_{k'})$ on the base change
$G_{k'}$ of $G$ to $k'$ by setting
\[
\phi_{k'} \ceq \phi \circ \Frob{G}^*(\phi) \circ \cdots \circ (\Frob{G}^{n-1})^*(\phi).
\]
The commutativity of the diagram (QC.3) for $\phi_{k'}$
follows from the fact that $\Frob{G_{k'}} = \Frob{G}^n$.
Note that we may also think about the construction of $\phi_{k'}$ from $\phi$
as taking the action $\varphi$ of $\Weil{k}$ on $\gqcs{L}$
defined in Section~\ref{sec:category} and restricting it to the subgroup $\Weil{k'}$.

@@

\begin{proof}
Let $\qcs{L} \ceq (\gqcs{L}, \mu, \phi) \in \QC(G)$ and write $F$ for $\Frob{G}$.  For any $x \in G(k')$,
we may compute the value of $t_{G_{k'}}(\nu_{k'/k}\qcs{L})(x)= t_{\nu_{k'/k}\qcs{L}}(x)$ as the trace of $\phi_{k'}$ on $\gqcs{L}_x$,
and the value of $\Nm^*(\TrFrob{G}(\qcs{L}))(x)$ as the trace of $\phi$ on $\gqcs{L}_{\Nm(x)}$.
Applying \ref{QC.3} to the stalk of $\gqcs{L}^{\boxtimes n}$ at the point $(x, \Frob{}(x), \ldots, \Frob{}^{n-1}(x))$ yields a diagram
\[
\begin{tikzcd}
\gqcs{L}_{\Nm(x)} \rar \dar{\phi_{\Nm(x)}} & \gqcs{L}_{F(x)} \otimes \gqcs{L}_{F^2(x)} \otimes \cdots \otimes \gqcs{L}_x
\dar{\phi_x \otimes (F^*\phi)_x \otimes \cdots \otimes ((F^{n-1})^*\phi)_x} \\
\gqcs{L}_{\Nm(x)} \rar & \gqcs{L}_x \otimes \gqcs{L}_{F(x)} \otimes \cdots \otimes \gqcs{L}_{F^{n-1}(x)}.
\end{tikzcd}
\]
Choose a basis vector $v$ for $\gqcs{L}_{\Nm(x)}$ and write $v_0 \otimes v_1 \otimes \cdots \otimes v_{n-1}$ for the image of $v$ under the
bottom map,
for $v_i \in \gqcs{L}_{\Frob{}^i(x)}$.  By \ref{QC.2}, $v$ maps to
$v_1 \otimes v_2 \otimes \cdots \otimes v_0$ along the top of the diagram.
Let $\alpha_i \in \EEx$ represent $((F^i)^*\phi)_x$ with respect to these bases and let $\alpha$ be
the trace of $\phi_{\Nm(x)}$.  We may now equate the trace $\alpha$ of $\phi$ on $\gqcs{L}_{\Nm(x)}$
with the product $\alpha_0 \cdots \alpha_{n-1}$, which is the trace of $\phi_{k'}$ on $\gqcs{L}_x$.
\end{proof}

@@

Let $K$, $R$ and $R_n$ be as above.
We will need the Greenberg functor, which provides
a functor from the category of schemes 
over $R_n$, for each $n\in \NN$, to the category of schemes over $\Fq$;
See \cite{bertrapelle-gonzales:Greenberg} for the definition and fundamental properties of the Greenberg functor as we use it; other important references include
\citelist{
\cite{greenberg:61}, \cite{greenberg:63a},
\cite{demazure-gabriel:GroupesAlgebriques}*{V, \S 4, no. 1},
\cite{bosch-lutkebohmert-reynaud:NeronModels}*{Ch. 9, \S 6},
\cite{nicaise-sebag:motivicSerre}*{\S 2.2}
}.
This functor respects open immersions, closed immersions, \'etale morphisms, smooth morphisms and geometric components.
%
For any non-negative integer $n$ we will write
\[
\Gr^R_n : \partop{\text{Schemes}}{\text{lft over $R$}} \to \partop{\text{Schemes}}{\text{lft over $\Fq$}}
\]
for the functor produced by composing pullback along $\Spec{R_n} \to \Spec{R}$ with the Greenberg functor. 
This functor also respects open immersions, closed immersions, \'etale morphisms, smooth morphisms and geometric components, and
\[
\Gr^R_n(X)(\Fq) \iso X(R_n),
\]
canonically, for any $X$ lft over $R$.

For any $n\leq m$,  the surjective ring homomorphism $R_{m} \to R_n$ determines a
natural transformation 
\[
\varrho^R_{n\leq m} : \Gr^R_{m} \to \Gr^R_n
\]
between additive functors.
Crucially, $\varrho^R_{n\leq m}(X): \Gr^R_{m}(X)\to \Gr^R_n(X)$ is an affine morphism of $\Fq$-schemes, for every $R$-scheme $X$ and every $n\leq m$ \cite{bertrapelle-gonzales:Greenberg}*{Prop 4.3}.
This observation is key to the proof that, for any $X$ lft over $R$, the projective limit 
\[
\Gr_R(X) \ceq \varprojlim_{n\in \NN} \Gr^R_n(X),
\]
taken with respect to the surjective morphisms $\varrho^R_{n\leq m}(X) : \Gr^R_{m}(X) \to \Gr^R_n(X)$,
exists in the category of group schemes over $\Fq$.
This leads to the definition of what we shall call the {\it Greenberg transform}:
\[
\Gr_R : \partop{\text{Schemes}}{\text{lft over $R$}} \to \partop{\text{Schemes}}{\text{lft over $\Fq$}}.
\]
By construction, the $\Fq$-scheme $\Gr_R(X)$ comes equipped with morphisms 
\[
\varrho^R_n(X) : \Gr^R(X) \to \Gr^R_n(X),\qquad \forall n\in \NN_0.
\]
The Greenberg transform respects open immersions, closed immersions, \'etale morphisms, smooth morphisms and geometric components, and
\[
\Gr_R(X)(\Fq) \iso X(R),
\]
canonically, for any $X$ lft over $R$.
